---
title: "Chicago Death Rate Predictors"
author: "Ebony Hunt"
date: "2026-03-12"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(AEDForecasting)
library(lattice)
library(tidyverse)
library(gamair)
library(mgcv)
```

# Aims

-   To ascertain the most important influences for predicting the death rates in Chicago, and how effective they are in predicting death rates.

-   Determine if air quality measures from 3 or 4 days beforehand predict the death rate better than those on the same day as the death rate.

-   To investigate if there simpler methods of prediction than the current process that are just as effective

# Background

The dataset "chicago" from the R package "gamair" (Peng, and Welty, 2004), contains information about daily non-accidental death rates in Chicago alongside measures of air quality and temperature from 1987 to 2000. 


# Findings


# Discussion

\newpage

# Statistical Methods and Results

## Computational Methods

All analysis was completed using R version 4.4.1 "Race for Your Life".

## Results

```{r warning=FALSE}
data("chicago")
chicago <- as.tibble(chicago)
head(chicago)

ggplot(data = chicago, mapping = aes(x = time, y = death)) +
  geom_point()

#looks like there is a seasonal component


ggplot(data = chicago, mapping = aes(x = tmpd, y = death)) +
  geom_point() +
  geom_smooth()

ggplot(data = chicago, mapping = aes(x = pm10median, y = death)) +
  geom_point() + 
  geom_smooth()


ggplot(data = chicago, mapping = aes(x = pm25median, y = death)) +
  geom_point() + 
  geom_smooth()

ggplot(data = chicago, mapping = aes(x = pm25median, y = death)) +
  geom_point() + 
  geom_smooth()

ggplot(data = chicago, mapping = aes(x = o3median, y = death)) +
  geom_point() + 
  geom_smooth()

ggplot(data = chicago, mapping = aes(x = so2median, y = death)) +
  geom_point() + 
  geom_smooth()

chicago
?chicago
summary(chicago)


```
